Refinement of Thermostated Molecular Dynamics via Backward Error Analysis

### 2018-06-10 rigid_bodies.tex
some corrections on disc. error

In the previous section we showed that the energy drift is mainly localized within the thermostat variables.
It is then reasonable to ask if this leads to a more ``accurate'' sampling of the physical system.
To answer this question, we turn our attention to the discretization errors.

In Fig.~\ref{fig:properties} we show the computed average of some thermodynamic properties as a function of $\timestep$. The simulations were carried out specifying T = 298 K and the analyzed quantities correspond to temperature ($T$), potential energy ($U$) and virial ($W$) per molecule, as well as the constant-volume heat capacity ($C_v$).
The lines are a guide for the eye. 
Also, the horizontal dotted lines (black) help to visualize the departure of the computed averages with increasing $\timestep$.
They correspond to the results obtained employing $\timestep$ = 1 fs for the numerical scheme proposed in this paper.
We have included the results of applying our method but using the approximate solution of rotation due to \citeauthor{Miller_2002} \cite{Miller_2002}, which is denoted as ``This work (NO-SQUISH)''.
The temperature values of Fig.~\ref{fig:properties} (a) were computed from the simulated kinetic energies assuming equipartition, ie $2 \langle K \rangle = (6N -3) k_b T$.

@@

From Fig.~\ref{fig:properties} it becomes clear that our method yields significantly improved results: we are able to simultaneously reproduce the specified temperature and the other thermodynamic properties. 
This is true also for the results corresponding to the use of the NO-SQUISH method\cite{Miller_2002} in the NVE part of the integrator.
Mathematically, the derivation of the refined Hamiltonian was facilitated by the fewer number of operators associated with the unsplit solution of rotation.
However, the accuracy of the approximate solution is very high for the time step sizes considered\cite{Silveira_2017} and this explains the good agreement.
Also note that a small, still clearly influencing systematic error is present in our scheme, as both temperature and potential energy are slightly modified by $\timestep$. This could, in principle, be remediated by using a higher-order approximation to the shadow Hamiltonian.

@@

In Fig.~\ref{fig:free_energy} we also depict the refined temperature ($\refined{T}$, squares), calculated by means of the generalized equipartition theorem, given by Eq.~\eqref{eq:equipartition}.
Of course, as $\refined{T}$ is indirectly controlled by means of Eq.~\eqref{eq:thermostat_force}, it is expected to reproduce the target temperature.
Indeed, for all the time-step sizes considered the agreement is excellent.

@@

In our previous work\cite{Silveira_2017}, we demonstrated that in the NVE case the rotational kinetic energy is more influenced by the time-step size because the rotational degrees of freedom are the fastest modes in the rigid model of water. 
We see that the method of \citeauthor{Kamberaj_2005} \cite{Kamberaj_2005} reproduces the behavior of the NVE case, as noted before.
The results of Fig.~\ref{fig:energy_partition} (b), which were obtained for the numerical scheme of \citeauthor{Martyna_1996} \cite{Martyna_1996}, are similar to Fig.~\ref{fig:energy_partition} (c), in the sense that the simulated translational and rotational kinetic energies deviate almost symmetrically and in opposite directions. Also, they do so contrary to what is expected: the translational kinetic energy should be smaller than the rotational one.
From Fig.~\ref{fig:energy_partition} it is clear that our method corrects that serious disruption for time-step sizes up to 5 fs.

>>>>

In the previous section we showed that the energy drift rate is mainly localized within the thermostat variables.
It is then reasonable to ask if this leads to a more accurate sampling of the physical system.
To answer this question, we turn our attention to the discretization errors.

In Fig.~\ref{fig:properties} we show the computed averages of some thermodynamic properties as a function of $\timestep$. The simulations were carried out specifying T = 298 K and the analyzed quantities correspond to temperature ($T$), potential energy ($U$) and virial ($W$) per molecule, as well as the constant-volume heat capacity ($C_V$).
The lines are a guide for the eye. 
Also, the horizontal dotted lines (black) help to visualize the departure of the computed averages with increasing $\timestep$.
They correspond to the results obtained employing $\timestep$ = 1 fs in the Refined/Unsplit numerical scheme.
We have included the results of applying our method but using the approximate solution of rotation due to \citeauthor{Miller_2002} \cite{Miller_2002}, which is denoted as ``Refined/NO-SQUISH''.
Recall that the computed averages for both Refined schemes are obtained via reweighting employing Eq.~\eqref{eq:reweighting}. 
In addition, it should be pointed out that the temperature values in Fig.~\ref{fig:properties} (a) were computed from the simulated kinetic energies assuming equipartition, ie $2 \langle K \rangle = (6N -3) k_b T$.

@@

From Fig.~\ref{fig:properties} it becomes clear that the Refined schemes yields significantly improved results: we are able to simultaneously reproduce the specified temperature and the other thermodynamic properties.
Note that a small, still clearly influencing systematic error is present, as the computed averages are slightly modified for $\timestep$ > 5.
This could, in principle, be remediated by using a higher-order approximation to the shadow Hamiltonian.
Mathematically, the derivation of the refined Hamiltonian was facilitated by the fewer number of operators associated with the unsplit solution of rotation.
However, the accuracy of the approximate solution is very high for the time step sizes considered\cite{Silveira_2017} and this explains the good agreement between schemes Refined/Unsplit and Refined/NO-SQUISH.

In Fig.~\ref{fig:properties} (a) we also depict the refined temperature, calculated by means of the generalized equipartition theorem, given by Eq.~\eqref{eq:equipartition}.
Of course, as $\refined{T}$ is indirectly controlled by means of Eq.~\eqref{eq:thermostat_force} we expect the Refined schemes to reproduce the target temperature.
Indeed, for all the time-step sizes considered the agreement is excellent.

@@

@@

In our previous work\cite{Silveira_2017}, we demonstrated that in the NVE case the rotational kinetic energy is more influenced by the time-step size because the rotational degrees of freedom are the fastest modes in the rigid model of water. 
We see that the method of \citeauthor{Kamberaj_2005} \cite{Kamberaj_2005} reproduces the behavior of the NVE case, as noted before.
The results of Fig.~\ref{fig:energy_partition} (b), which were obtained for the numerical scheme of \citeauthor{Martyna_1996} \cite{Martyna_1996}, are similar to Fig.~\ref{fig:energy_partition} (c), in the sense that the simulated translational and rotational kinetic energies deviate almost symmetrically and in opposite directions.
Also, they do so contrary to what is expected: the translational kinetic energy should be smaller than the rotational one.
Fig.~\ref{fig:energy_partition} (a) was obtained directly from the contribution of translational and rotational degrees of freedom to the refined kinetic energy.
It confirms that our approximation bla bla bla
After reweighting, which yields Fig.~\ref{fig:energy_partition} (b), it becomes clear that our method corrects, for time-step sizes up to 5 fs, the serious disruption observed in the schemes of \citeauthor{Martyna_1996} \cite{Martyna_1996} and  \citeauthor{Kamberaj_2005} \cite{Kamberaj_2005}
